### 2014-04-01 paper/abstract.tex
shortening

\begin{abstract}
3D printers offer extraordinary flexibilty in prototyping the shape and mechanical function of objects.  However, a 3D printer's role in prototyping interactive objects is still not clear.  We propose a general technique for adding interactivity by removing interior material to form pipes and cavities. These can be used to sense input and generate haptic as well as visible output. We describe the design space of pipes and hollow chambers for interaction design: a variety of openings, topologies, and inserted media for pipes create diverse inputs and outputs.  We present a technique and design tool for routing pipes of various topologies through the interior of 3D printed parts.  Our design tool is integrated into a 3D model manipulation program.  We use two distinct routing algorithms.  One, for creating particular points of interaction or integrating with existing electronics, allows users to select begin and end points for pipes, then uses A* path routing and physics-based simulation to minimize the bending energy of routed paths.  The second, for creating, e.g., novel neon signs, allows users to enter a description of paths to follow: for this we developed a novel routing algorithm and offer proof of key traits.  We present several totally tubular prototypes we created using this tool to show its flexibility and potential, as well as to explore new points in the pipe design space.
\end{abstract}

>>>>

\begin{abstract}
3D printers offer extraordinary flexibilty in prototyping the shape and mechanical function of objects.  However, a 3D printer's role in prototyping interactive objects is still not clear.  We propose a general technique for adding interactivity by removing interior material to form pipes and cavities. We describe the design space of pipes and hollow chambers for interaction design, touching on of openings, topologies, and inserted media.  We present a technique and design tool for routing pipes through the interior of 3D models.  We integrated our prototype with a 3D modeling program.  We use two distinct routing algorithms.  One, for creating particular points of interaction or integrating with existing electronics on the model's surface, allows users to select pipes' terminals, then uses A* path routing and physics-based simulation to minimize the bending energy of routed paths.  The second, for creating, e.g., novel neon signs, has users supply a desired internal shape which we fit a pipe route to: for this we developed a novel graph-based routing algorithm.  We present several totally tubular prototypes we created using our tool to show its flexibility and potential.
\end{abstract}

### 2014-04-01 paper/designspace.tex
shortening

 There still some redundancy with the related work section, too}
By using different types and topologies of pipes and inserting different media into the completed pipes, a variety of inputs and outputs can be located at arbitrary points on the surface of a 3D print.  We describe the space below.

\begin{figure}[h!]
\centering
    \includegraphics[width=3.4in]{figures/tubespace.pdf}
\caption{The design space of pipes.  Pipe types, media, topologies, and design are discussed more fully in the text.}
\label{fig:pipespace}
\end{figure}

@@

\tovi{this seems like a forced and
unnecessary definition. Do we really
need to distibguish the two ends?}
Pipes can come in four types: open, return, semi-closed, and fully enclosed.  Each of these types offers distinct interaction capabilities (see Figure \ref{fig:pipespace}).  Since we are interested in interactive devices, we distinguish a ``user side" - part of the surface or interior of an object facing its user; and a ``system side" - part of the object (either interior or exterior) where other hardware such as electronics will be connected.  Definitionally, in the following paragraphs we refer to a pipe which is ``open on the user side'' or ``open on the system side'' to have its contents \emph{physically accessible} to an interacting user or the system \emph{at interaction time}.  A pipe which is ``open on the user side'' or ``closed on the system side'' is physically cut off from the user or machine.  User and system manipulation of contents are possible with or without physical access, e.g., by resonating, pressurizing, or passing electricity through those contents.

@@

Open pipes (see Figure \ref{fig:pipespace}A) originate from the system side and connect the user side, with both ends of the pipe open. This type of pipe may be used to create, for example, capacitive sensors: an open pipe filled with conductive paint can be connected to a sensing platform (e.g., Arduino) on the system side, while a user can touch the uncapped other end of the pipe.  Using Swept Frequency Capacitive Sensing \cite{Sato-touche} or other techniques, a user's touch of the open end can be sensed.  An open pipe can also be used for output, for example by creating in-air vortices as in \cite{Sodhi-aireal}.

A return pipe (see Figure \ref{fig:pipespace}B) originates at the system and returns back to it, without offering physical access to the user.  By threading an electroluminescent (EL) wire through a clear return pipe, a maker can create a custom piece of neon art.  If a return pipe passes very close to the surface of a 3D printed model, warm or cold water passed through the pipe could be used for temperature-based haptic feedback.

Semi-closed pipes (see Figure \ref{fig:pipespace}C) are open at the system end (for control of the enclosed medium) and closed at the user end.  The closed interface  may be fabricated from a malleable material: for example, a series of pipes terminated in thin rubber membranes on the user side can be actuated by an air pump to create haptic feedback, or a hinged cap can be blown open by air pressure.  Semi-closed pipes can also function as audio-generating resonance chambers.

A fully enclosed pipe (see Figure \ref{fig:pipespace}D) is closed to both the system side and the user side.  Fully enclosed pipes can be used as resonance chambers (e.g., for object identification), or as air bubbles (e.g., as used in Printed Optics (\cite{Willis-printedoptics}) for internal display).  They can also be used as containers for water or other particles that would otherwise fall out.

\subsection{Design of Pipes}
\tovi{This definition is a little bit vague.}
Pipe designs may emphasize either their exterior features (connection points) or their interior paths.  These two features lead to different kinds of interfaces.  An example interface that focuses on the exterior connection points is the touch sensitive toy in Figure \ref{fig:toys}, where pipes must exit the toy at the frontal lobe, parietal lobes, etc.  An interface focused on the internal path of the pipe is the neon sign in Figure \ref{fig:neon}: output is based on the shape of the pipe.

@@

Pipe network topologies enable different types of interactions.

>>>>

 There still some redundancy with the related work section, too}
By using different openings, designs, and topologies of pipes, and inserting different media into the completed pipes, inputs and outputs can be located at arbitrary points on the surface of a 3D print.  We describe the space below.

\begin{figure}[h!]
\centering
    \includegraphics[width=3.4in]{figures/tubespace.pdf}
\caption{The design space of pipes.  Pipe openings, design, topologies, and media are discussed more fully in the text.}
\label{fig:pipespace}
\end{figure}

@@

\tovi{this seems like a forced and
unnecessary definition. Do we really
need to distibguish the two ends?}
Pipes can have four opening styles: open, return, semi-closed, and fully enclosed.  Each of these types offers distinct interaction capabilities (see Figure \ref{fig:pipespace}).  As we are interested in interactive devices, we distinguish a ``user side"---part of the surface or interior of an object facing its user; and a ``system side"---where other hardware such as electronics may be connected.  Definitionally, in the following paragraphs we refer to a pipe which is ``open on the user side'' or ``open on the system side'' to have its contents \emph{physically accessible} to an interacting user or the system \emph{at interaction time}.  Similarly, pipes ``closed on the user side'' or ``closed on the system side'' are physically cut off from the user or machine.  User and system manipulation of contents are possible with or without physical access, e.g., by resonating, pressurizing, or passing electricity through those contents.

@@

Open pipes originate from the system side and connect the user side, with both ends of the pipe open. This type of pipe may be used to create electronic circuitry.  An open pipe filled with conductive paint can power an LED (see Figure \ref{fig:pipespace}A).

A return pipe originates at the system and returns back to it, without offering physical access to the user.  By threading an electroluminescent (EL) wire through a clear return pipe, a maker can create a custom piece of neon art (see Figure \ref{fig:pipespace}B).

Semi-closed pipes are open at the system end and closed at the user end.  The closed interface may be fabricated from a malleable material: for example, a series of pipes terminated in thin rubber membranes on the user side can be actuated by an air pump to create haptic feedback (see Figure \ref{fig:pipespace}C).

A fully enclosed pipe is closed to both the system and the user.  Fully enclosed pipes may be used as weight modifiers (e.g., for object identification), as in Figure \ref{fig:pipespace}D).  They can also be used as containers for water or other particles that would otherwise fall out.

\subsection{Design of Pipes}
Pipe design may emphasize either exterior connection points or interior paths.  These two foci lead to different kinds of interfaces.  An example focused on the exterior features is the touch sensitive toy in Figure \ref{fig:toys}, where pipes must exit the toy at the frontal lobe and parietal lobes to make those \emph{specific locations} touch-sensitive.  Focused on the internal features is the neon sign in Figure \ref{fig:neon}: the sign's output is based on the \emph{path} of the pipe.

@@

Pipe network topologies enable different interactions.
